---
title: "amb_analysis"
author: "maria"
date: "2024-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,
               leaflet) 

# SETTING SEED FOR REPRODUCABILITY
set.seed(42)
```

# LOADING IN DATA
```{r}
scaled_data <- read_csv("scaled_data.csv")
all_events <- read_csv("all_events.csv")
matched_people <- read_csv("matched_people.csv")
population_df <- read.csv("population_df.csv")

# CONVERTING COLUMNS TO LISTS OF CHARACTER VECTORS 
scaled_data$neighbors <- strsplit(scaled_data$neighbors, ", ")
scaled_data$event_category <- strsplit(scaled_data$event_category, ", ")

matched_people$neighbors <- strsplit(matched_people$neighbors, ", ")
matched_people$event_category <- strsplit(matched_people$event_category, ", ")
matched_people$attending_with <- strsplit(matched_people$attending_with, ", ")
matched_people$attending_events <- strsplit(matched_people$attending_events, ", ")

```



# INTERACTIVE MAP OF PEOPLE LOCATION
```{r}
#PLOTTING
people_map <- leaflet() %>%
  addTiles() %>%
  setView(lng = 10.5, lat = 56, zoom = 7) 

people_map <- people_map %>%
  addMarkers(data = scaled_data, lng = ~Longitude, lat = ~Latitude,
             popup = ~paste("ID:", ID, "<br>",
                            "Age:", age, "<br>",
                            "Gender:", gender, "<br>",
                            "Intressts:", event_category, "<br>"))



people_map
```


# HEATMAP OF PEOPLE LOCATION
```{r}
heatmap_people <- leaflet(scaled_data) %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addHeatmap(
    lng = ~Longitude,
    lat = ~Latitude,
    intensity = ~1,  
    blur = 20,       
    max = 0.05,      
    radius = 15      
  ) %>%
  setView(lng = 10.0, lat = 56.0, zoom = 6)

heatmap_people
```


# INTERACTIVE MAPE OF EVENT LOCATION
```{r}
event_map <- leaflet() %>%
  addTiles() %>%
  setView(lng = 10.5, lat = 56, zoom = 7) 

event_map <- event_map %>%
  addMarkers(data = all_events, lng = ~Longitude, lat = ~Latitude,
             popup = ~paste("Event:", Event_Name, "<br>",
                            "Date:", Event_date, "<br>"))

event_map
```


# HEATMAP OF EVENT LOCATION
```{r}
heatmap <- leaflet(all_events) %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addHeatmap(
    lng = ~Longitude,
    lat = ~Latitude,
    intensity = ~1,  # Optional, could be based on some attribute
    blur = 20,       # Optional, controls the blur effect
    max = 0.05,      # Optional, controls the intensity of the heatmap
    radius = 15      # Optional, controls the radius of the heatmap
  ) %>%
  setView(lng = 10.0, lat = 56.0, zoom = 6)

heatmap
```


# INTERACTIVE MAP OF ATTENDED EVENTS
```{r}
# FINDING THE ATTENDED EVENTS BASED ON ID 
event_ids <- unique(unlist(matched_people$attending_events))

attended_events <- events_data %>%
  filter(Event_ID %in% event_ids)


# MAPPING ATTENDED EVENTS 
attended_event_map <- leaflet() %>%
  addTiles() %>%
  setView(lng = 10.5, lat = 56, zoom = 7) 

attended_event_map <- attended_event_map %>%
  addMarkers(data = attended_events, lng = ~Longitude, lat = ~Latitude,
             popup = ~paste("Event:", Event_Name, "<br>",
                            "Date:", Event_date, "<br>"))

attended_event_map
```


# HEATMAP OF ATTENDED EVENTS
```{r}
# HEATMAP OF ATTENDED EVENTS 
attended_heatmap <- leaflet(attended_events) %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addHeatmap(
    lng = ~Longitude,
    lat = ~Latitude,
    intensity = ~1,  
    blur = 20,       
    max = 0.05,      
    radius = 15      
  ) %>%
  setView(lng = 10.0, lat = 56.0, zoom = 6)

attended_heatmap  
```


# CALCULATING DISTANCE (IN KM) FROM PERSON TO EVENT
```{r}
# MAKING NEW COLUMN 
matched_people$distance_to_event <- vector("list", nrow(matched_people))

# USING THE HAVERSINE FORMULA TO CALCULATE THE GREAT-CIRCLE DISTANCE BETWEEN TWO POINTS, WHICH IS THE SHORTEST DISTANCE OVER THE EARTH'S SURFACE
for (i in 1:nrow(matched_people)) {
  distances <- c()
  
  for (j in 1:length(matched_people$attending_events[[i]])) {
    lon1 <- matched_people$Longitude[i]
    lat1 <- matched_people$Latitude[i]
    
    event_id <- matched_people$attending_events[[i]][j]
    event_id_index <- which(attended_events$Event_ID == event_id)

    if (length(event_id_index) > 0) {
      lon2 <- attended_events$Longitude[event_id_index]
      lat2 <- attended_events$Latitude[event_id_index]
      
      # CALCULATING DISTANCE
      distance <- distHaversine(c(lon1, lat1), c(lon2, lat2))
      
      # CONVERTING DISTANCE FROM METERS TO KILOMETERS 
      distance_km <- distance / 1000
      
      # ADDING DATA 
      distances <- c(distances, distance_km)
    }
  }
  matched_people$distance_to_event[[i]] <- distances
}

# CALCULATING THE OVERALL TRAVEL MEAN 
all_distances <- unlist(matched_people$distance_to_event)

# CALCULATING THE MEAN OF ALL DISTANCES 
mean_distance <- mean(all_distances)
sd_distance <- sd(all_distances)

# CALCULATING MEAN TRAVEL DISTANCE 
matched_people$mean_travel_dist <- NA
matched_people$sd_travel_dist <- NA

for (i in 1:nrow(matched_people)) {
  matched_people$mean_travel_dist[i] <- round(mean(matched_people$distance_to_event[[i]]), 2)
  matched_people$sd_travel_dist[i] <- round(sd(matched_people$distance_to_event[[i]]), 2)
}

# FINDING MINIMUM AND MAXIMUM MEAN TRAVEL DISTANCE 
min_mean_travel_dis <- min(matched_people$mean_travel_dist)
max_mean_travel_dis <- max(matched_people$mean_travel_dist)
```



# SEEING IF PEOPLE FROM SMALLER CITIES ARE WILLING TO TRAVEL FURTHER FOR EVENTS 
```{r}
# MERGING POPULATION DATA WITH PEOPLE DATA 
matched_people <- merge(matched_people, population_df, by = "Municipality")

distance_data <- matched_people %>%
  unnest(distance_to_event) %>%  # Flatten the list column
  group_by(Municipality) %>%
  summarise(Average_Distance = round(mean(distance_to_event), 2), Population = first(Population))

# COMPARING DISTANCES BASED ON POPULATION SIZE 
# FOR SIMPLICITY, MUNICIPALITIES ARE DIVIDED INTO SMALL AND LARGE BASED ON MEDIAN POPULATION 
distance_data <- distance_data %>% 
  mutate(Population = as.numeric(Population))

population_df <- population_df %>% 
  mutate(Population = as.numeric(Population))

median_population <- median(population_df$Population)

distance_data <- distance_data %>%
  mutate(Size_Category = ifelse(Population < median_population, "Small", "Large"))




# PLOTTING  
travel_distance_plot_box <- ggplot(distance_data, aes(x = Size_Category, y = Average_Distance)) +
  geom_boxplot() +
  labs(title = "Travel Distance by Municipality Size", x = "Municipality Size", y = "Average Travel Distance (km)") +
  theme_minimal()

travel_distance_plot_size <- ggplot(distance_data, aes(x = Size_Category, y = Average_Distance)) +
  geom_col(fill = "pink") +
  labs(title = "Travel Distance by Municipality Size", x = "Municipality Size", y = "Average Travel Distance (km)") +
  theme_minimal()

travel_distance_plot_city <- ggplot(distance_data, aes(x = Municipality, y = Average_Distance)) +
  geom_col(fill = "pink") +
  labs(title = "Travel Distance by Municipality", x = "Municipality", y = "Average Travel Distance (km)") +
  theme_minimal()

print(travel_distance_plot_box)
print(travel_distance_plot_size)
print(travel_distance_plot_city)
```


# FINDING TOTAL NUMBER OF MATCHES
```{r}
# FLATTENING THE LISTS AND CREATING PAIRS 
pairs <- matched_people %>%
  unnest(attending_with) %>%
  rename(person1 = ID, person2 = attending_with)

# CREATING PAIRS IN BOTH DIRECTIONS 
all_pairs <- pairs %>%
  bind_rows(pairs %>% rename(person1 = person2, person2 = person1))

# IDENTIFYING MUTUAL MATCHES 
mutual_matches <- all_pairs %>%
  semi_join(all_pairs, by = c("person1" = "person2", "person2" = "person1")) %>%
  distinct(person1, person2)

# COUNTING NUMBER OF UNIQUE MATCHES 
# EACH MUTUAL PAIR WILL APPEAR TWICE, SO DEVINDING BY 2 TO GET ACCUAL COUNT 
number_of_matches <- nrow(mutual_matches) / 2

```


