---
title: "simulating_data"
author: "maria"
date: "2024-06-05"
output: html_document
---

link to denmark municipality data: https://gadm.org/download_country.html (gadm41_DNK_2.shp)
link to population data: https://www.dst.dk/da/Statistik/emner/borgere/befolkning/befolkningstal (FOLK1AM.csv)
link to information that city code ranges data is based on: https://edemann.dk/liste-over-kommuner-og-postnumre/ (but made myself, Municipality_city_codes.csv)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,
               sf) 

# SETTING SEED FOR REPRODUCABILITY
set.seed(42)
```


#CLEANING SURVEY DATA

```{r}
# LOADING IN SURVEY DATA
people_data <- read_csv("survey_data.csv")

# RENAMING COLUMNS
people_data_cleaned <- people_data %>% 
  rename(age = `What is your age?`,
         gender = `What is your gender?`,
         consent = `I have read and understood the above consent statement and agree to participate.`,
         city = `In which city do you live in? (answer in postal code)`,
         event_type = `What types of events do you enjoy attending? (Select all that apply)`,
         event_category = `Which of the following categories interesst you? (Select all that apply)`,
         event_freq = `How often do you attend events?`,
         travel_willingness = `How far are you willing to travel for an event?`,
         importance_meet_people = `How important is it for you to meet new people at events?`,
         age_connect = `What age are the people you typically connect with?`,
         gender_connect = `What gender are the people you typically connect with? (Select all that apply)`,
         maintain_contact = `How often do you maintain contact with new people you connect with at events?`,
         wish_maintain_contact = `How often do you wish you maintained contact with new people you connected with at events?`) %>% 

  # FILTERING OUT COLUMNS
  select(-"Tidsstempel", -"consent", -"event_type") 

# REMOVING COLUMNS WHERE PEOPLE DIDN'T ANSWER ALL QUESTIONS
people_data_cleaned <- na.omit(people_data_cleaned)





# MODIFYING THE "INTRESTS" COLUMN TO CONTAIN LISTS OF EACH PEOPLES INTRESTS
people_data_cleaned$event_list <- vector("list", nrow(people_data_cleaned))
for (i in 1:nrow(people_data_cleaned)) {
  intrests <- unlist(strsplit(people_data_cleaned$event_category[i], ", "))
  for (j in 1:length(intrests)) {
    people_data_cleaned$event_list[[i]][j] <- intrests[j]
  }
}

# FUNCTION TO CORRECT THE "CARS & BOATS & PLANES" INTREST VALUE
merge_event_categories <- function(event_list) {
  event_list <- map(event_list, function(x) {
    if ("Cars" %in% x && "Boats & Planes" %in% x) {
      x <- x[!(x %in% c("Cars", "Boats & Planes"))]  # Remove "Cars" and "Boats & Planes"
      x <- c(x, "Cars & Boats & Planes")  # Add the combined category
    } else if ("Cars" %in% x) {
      x <- str_replace_all(x, "Cars", "Cars & Boats & Planes")
    } else if ("Boats & Planes" %in% x) {
      x <- str_replace_all(x, "Boats & Planes", "Cars & Boats & Planes")
    }
    return(unique(x))  # Ensure each category is unique
  })
  return(event_list)
}

# CORRECTING THE "CARS & BOATS & PLANES" INTREST VALUE
people_data_cleaned <- people_data_cleaned %>%
  mutate(event_list = merge_event_categories(event_list)) %>% 
  select(-event_category) %>% 
  rename(event_category = event_list)



# CORRECTING COLUMN CLASSES
 people_data_cleaned <- people_data_cleaned %>% 
   mutate(
    gender = as.factor(gender),
    city = as.numeric(city),
    event_freq = as.factor(event_freq),
    travel_willingness = as.factor(travel_willingness),
    importance_meet_people = as.factor(importance_meet_people),
    maintain_contact = as.factor(maintain_contact),
    wish_maintain_contact = as.factor(wish_maintain_contact),
    age_connect = as.factor(age_connect)
  )


# RECODING THE AGE_CONNECT AND TRAVEL_WILLINGNESS COLUMNS
people_data_cleaned <- people_data_cleaned %>%
  mutate(age_connect = recode(age_connect,
                              "Same age as myself" = "same",
                              "Younger than myself" = "younger",
                              "Older than myself" = "older")) %>% 
  
  mutate(travel_willingness = recode(travel_willingness,
                                    "To neighboring cities" = "neighbouring",
                                    "Anywhere in Denmark" = "anywhere",
                                    "Within my city" = "in_city",
                                    "Within the area i live in (Sjælland, Jylland, Fyn, Bornholm)" = "region"))



# MAKING ALL CHARECTERS TO LOWER CASE
names(people_data_cleaned) <- tolower(names(people_data_cleaned))

people_data_cleaned <- people_data_cleaned %>%
  mutate(across(c(gender, event_freq, 
                  gender_connect, maintain_contact, wish_maintain_contact), tolower)) 



```

# MAKING A DATAFRAME OF INFORMATION ON MUNICIPALITIES AND POPULATION SIZE
```{r}
# LOADING IN SPATIAL DATA ON DENMARKS MUNICIPALITIES
DK_border <- st_read("gadm41_DNK_2.shp")

# LOADING DATA ON MUNICIPALITY POPULATION SIZE
population_df <- read.csv("FOLK1AM.csv")

# CLEANING POPULATION DATA
population_df <- population_df %>%
  separate(OMRÅDE.TID.INDHOLD, into = c("Municipality", "Date", "Population"), sep = ";") %>% 
  # ONLY KEEPING NEWEST DATA (FROM APRIL 2024) 
  filter(Date == "2024M04") %>% 
  # FILTERING OUT REGIONAL POPULATION DATA 
  filter(Municipality!="Hele landet",
         Municipality!= "Region Hovedstaden",
         Municipality!= "Region Syddanmark",
         Municipality!= "Region Sjælland",
         Municipality!= "Region Midtjylland",
         Municipality!= "Region Nordjylland") %>%
  # RECODING SOME MUNICIPALITY NAMES (SO COMPATIBLE WITH DK_BORDER DATA)
  mutate(Municipality = recode(Municipality,
                              "Aarhus" = "Århus",
                              "Høje-Taastrup" = "Høje Taastrup",
                              "Vesthimmerlands" = "Vesthimmerland"))

# MERGING THE TWO DATAFRAMES TO MAKE ONE DATAFRAME WITH SCATIAL DATA AND POPULATION SIZE OF MUNICIPALITIES IN DENMARK
DK_border <- merge(population_df, DK_border, by.x = "Municipality", by.y = "NAME_2", all.x = TRUE)

# MAKING SURE POPULATION COLUMN IS NUMERIC
DK_border <- DK_border %>% 
  mutate(Population = as.numeric(Population))
```


# SIMULATING EVENT DATA
```{r}

# CALCULATING MUNICIPALITY NUMBER OF EVENTS BASED ON POPULATION SIZE
scaling_factor <- 0.0001
DK_border$Event_Count <- ceiling(DK_border$Population * scaling_factor)

# FUNCTION TO GENERATE RANDOM POINTS WITHIN A POLYGON
generate_random_location <- function(multipolygon, n) {
  points <- st_sample(multipolygon, size = n)
  coords <- st_coordinates(points)
  return(data.frame(Latitude = coords[, 2], Longitude = coords[, 1]))
}

# FUNCTION TO GENERATE EVENTS FOR A MUNICIPALITY
generate_events <- function(city_row, multipolygon) {
  # GET A RANDOM DATE IN THE MONTH OF JUNE 2024
  event_date <- as.POSIXct("2024-06-01") + days(sample(0:29, 1))
  
  # GENERATING RANDOM LOCATIONS WITHING THE MULTIPOLYGON
  locations <- generate_random_location(multipolygon, city_row$Event_Count)
  
  # MAKING AN EVENTS DATAFRAME WITH A AN EVENT NAME THAT REPRECENTS THE CATEGORY OF AN EVENT
  events <- data.frame(
    Event_Name = sample(c("Music", "Art", "Science & Tech", "Food & Drink", "Health", "Business", "Cars & Boats & Planes", "Charity", "Education", "Film & Media", "Performing and Visual Arts", "Spirituality", "Sports & Fitness", "Travel & Outdoor"), city_row$Event_Count, replace = TRUE),
    Event_date = event_date,
    Latitude = locations$Latitude,
    Longitude = locations$Longitude,
    Venue_Name = city_row$Municipality
  )
  return(events)
}


# GENERATING EVENTS FOR EACH MUNICIPALITY
all_events <- bind_rows(lapply(1:nrow(DK_border), function(i) {
  generate_events(DK_border[i, ], DK_border$geometry[i])
}))

# GENERATING UNIQUE EVENT ID
all_events <- all_events %>%
  mutate(Event_ID = paste0("event_", row_number()))

# ADDING A REAGION COLUMN WITH THE REAGION IN DENMARK THAT THE EVENTS TAKES PLACE IN 
all_events <- all_events %>%
  left_join(DK_border %>% select(Municipality, NAME_1), by = c("Venue_Name" = "Municipality")) %>%
  rename(Region = NAME_1) 

# MAKING SURE THE DATE COLUMNS IS CLASS = DATE
all_events <- all_events %>% 
  mutate(Event_date = as.Date(Event_date))
```

# ADDING MUNICIPALITY INFORMATION AND NEIGHBOURING MUNICIPALITY INFORMATION TO THE CLEANED SURVEY DATA
```{r}
# LOADING IN DATA ON MUNICIPALITY NAMES AND CITY CODES
Municipality_code_ranges <- read_csv("Municipality_city_codes.csv")

# CLEANING MUNICIPALITY_CODE_RANGES
Municipality_code_ranges <- Municipality_code_ranges %>%
  separate(`Municipality;city_codes_min;city_codes_max;;;;;;;`, into = c("Municipality", "city_codes_min", "city_codes_max"), sep = ";") %>% 
  mutate(city_codes_min = as.numeric(city_codes_min),
         city_codes_max = as.numeric(city_codes_max))
  
# REPLACING "OE" WITH "Ø"
Municipality_code_ranges$Municipality <- gsub("oe", "ø", Municipality_code_ranges$Municipality)

# REPLACING "AE" WITH "Æ" 
Municipality_code_ranges$Municipality <- gsub("ae", "æ", Municipality_code_ranges$Municipality)

# REPLACING "AAA" WITH "Å" 
Municipality_code_ranges$Municipality <- gsub("aaa", "å", Municipality_code_ranges$Municipality)

# RECODING SOME MUNICIPALITY NAMES FOR COMPATABILITY
Municipality_code_ranges <- Municipality_code_ranges %>%
  mutate(Municipality = recode(Municipality,
                              "århus" = "Århus",
                              "ærø" = "Ærø"))



### CATEGORIZING WHICH MUNICIPALITY PEOPLE LIVE IN BASED ON THEIR POSTAL CODES

# MAKING SURE THERE ARE NO EMPTY VALUES
people_data_cleaned <- na.omit(people_data_cleaned)

# INITIALIZING MUNICIPALITY COLUMN 
people_data_cleaned$Municipality <- NA



# LOOPING THROUGH ALL ROWS IN PEOPLE_DATA_CLEANED
for (i in 1:nrow(people_data_cleaned)) {
  # GETTING THE CITY CODE FOR THE CURRENT PERSON
  city_code <- people_data_cleaned$city[i]
  
  # INITIALIZING MINICIPALITY COLUMN 
  municipality <- NA
  
  # CHECKING WHICH MUNICIPALITY THE CITY CODE BELONGS TO
  for (j in 1:nrow(Municipality_code_ranges)) {
    if (city_code >= Municipality_code_ranges$city_codes_min[j] & 
        city_code <= Municipality_code_ranges$city_codes_max[j]) {
      municipality <- Municipality_code_ranges$Municipality[j]
      break  # Exit the loop once a match is found
    }
  }
  
  # ASSIGNING THE MUNICIPALITY (OR NA) TO THE CURRENT ROW
  people_data_cleaned$Municipality[i] <- municipality
}


# REMOVING EMTY VALUES
people_data_cleaned <- na.omit(people_data_cleaned)





# ADDING A REGION COLUMN 
people_data_cleaned <- people_data_cleaned %>%
  left_join(DK_border %>% select(Municipality, NAME_1), by = c("Municipality" = "Municipality")) %>%
  rename(region = NAME_1)



# ADDING A NEIGHBORING MUNICIPALITIES COLUMN

# LOADING IN SPATIAL DATA ON DENMARKS MUNICIPALITIES AGAIN
DK_border_2 <- st_read("gadm41_DNK_2.shp")

# CONVERTING DK_BORDER TO AN SF OBJECT IF NOT ALREADY
if (!inherits(DK_border_2, "sf")) {
  DK_border_2 <- st_as_sf(DK_border_2, wkt = "geometry")
}

# CREATING A LIST OF NEIGHBORING MUNICIPALITIES 
neighbor_list <- st_touches(DK_border_2)

# ADDING THE NEIGHBORING MUNICIPALITIES TO THE DK_BORDER DATAFRAME
DK_border_2$neighbors <- lapply(neighbor_list, function(nbrs) {
  DK_border_2$NAME_2[nbrs]
})

# CREATING A VECTOR OF MUNICIPALITIY NEIGHBOURS FOR EASIER LOOKUP 
municipality_neighbors <- setNames(DK_border_2$neighbors, DK_border_2$NAME_2)

# ADD NEIGHBORS TO PEOPLE_DATA_CLEANED
people_data_cleaned <- people_data_cleaned %>%
  rowwise() %>%
  mutate(neighbors = list(municipality_neighbors[[Municipality]]))
```

# SCALING SURVEY DATA (N= 200)
```{r}
# FUNCTION FOR SCALING DATA USING BOOTSTRAPPING 
scale_data <- function(data, target_size) {
  n <- nrow(data)
  indices <- sample(1:n, target_size, replace = TRUE)
  scaled_data <- data[indices, ]
  return(scaled_data)
}

# SCALLING DATA TO 200 SAMPLES (CHANGE NUMBER HERE IF WISHED TO DO ANALYSIS ON A LARGER SAMPLE SIZE) 
scaled_data <- scale_data(people_data_cleaned, 500)

# ENSURING UNIQUE IDS
all_ids <- 1:nrow(scaled_data)

for (i in 1:nrow(scaled_data)) {
  scaled_data$ID[i] <- paste0("PER_",all_ids[i]) 
}
```

# SIMULATING/ASSIGNING A LOCATION FOR EACH PERSON IN SCALED DATA
```{r}

# CREATING A NAMED LIST OF GEOMETRIES INDEXED BY MUNICIPALITY NAMES
municipality_geometries <- split(DK_border_2$geometry, DK_border_2$NAME_2)

# INITIALIZING COLUMNS FOR LATITUDE AND LONGITUDE IN PEOPLE_DATA_CLEANED
scaled_data$Latitude <- NA
scaled_data$Longitude <- NA

# FUNCTION TO GENERATE RANDOM LOCATION WITHIN A POLYGON
generate_random_location <- function(multipolygon, n) {
  points <- st_sample(multipolygon, size = n)
  coords <- st_coordinates(points)
  return(data.frame(Latitude = coords[, 2], Longitude = coords[, 1]))
}

# LOOPING THROUGH EACH PERSON AND GENERATING RANDOM A LOCATION WITHIN THEIR MUNICIPALITY

for (i in 1:nrow(scaled_data)) {
  municipality_name <- scaled_data$Municipality[i]
  municipality_geom <- municipality_geometries[[municipality_name]]
  if (!is.null(municipality_geom) && !is.na(municipality_geom)) {
    random_location <- generate_random_location(municipality_geom, 1)
    scaled_data$Latitude[i] <- random_location$Latitude
    scaled_data$Longitude[i] <- random_location$Longitude
  }
}
```

# SAVING RELEVANT DATA AS CSV FILES IN CURRENT DICTIONARY
```{r}
# CONVERTING THE LIST COLUMNS INTO STRINGS SO IT CAN BE SAVED AS A CSV FILE
scaled_data$neighbors <- sapply(scaled_data$neighbors, toString)
scaled_data$event_category <- sapply(scaled_data$event_category, toString)


# SAVING TO CSV FILES 
write.csv(scaled_data, "scaled_data.csv", row.names = FALSE)
write.csv(all_events, "all_events.csv", row.names = FALSE)
write.csv(population_df, "population_df.csv", row.names = FALSE)

```

